Bayesian multi-proxy reconstruction of early Eocene latitudinal temperature gradients

### 2023-05-31 manuscript/main.qmd
Update main.qmd

Our results show that relatively small (*n* = 10 - 20) and patchy samples of temperature data are sufficient to reconstruct greenhouse- and icehouse-type, latitudinal SST gradients, using our Bayesian model. This is an advancement over previously used linear, quadratic or Gaussian approximations [e.g. @Bijl2009; @Tierney2017], which can fit only specific types of gradients. As such, our model presents an alternative to non-parametric methods for inferring latitudinal temperature gradients, which are sometimes favoured as they can very flexibly follow the shape of an unknown temperature gradient [e.g. @Zhang2019; @Jones2022]. However, when used for interpolation or prediction outside the proxy range, non-parametric methods such as Gaussian process regression strictly respond to the data [e.g. @Inglis2020]. This means that the idiosyncrasies of a patchy proxy record, potentially afflicted with measurement errors, calibration errors, and palaeogeographic and temporal uncertainty, directly generates the resulting large-scale climate patterns.

>>>>

Our results show that our Bayesian model can be used to reconstruct different types of latitudinal SST gradients from proxy data, even with small sample sizes (*n* = 10 - 20) and patchy sampling distributions (Fig. 2). This is an advancement over previously used linear, quadratic or Gaussian approximations [e.g. @Bijl2009; @Tierney2017], which can fit only specific types of gradients. As such, our model presents an alternative to non-parametric methods for inferring latitudinal temperature gradients, which are sometimes favoured as they can flexibly follow the shape of an unknown temperature gradient [e.g. @Zhang2019; @Jones2022]. However, when used for interpolation or prediction outside the proxy range, non-parametric methods such as Gaussian process regression strictly respond to the data [e.g. @Inglis2020]. This means that the idiosyncrasies of a patchy proxy record, potentially afflicted with measurement errors, calibration errors, and palaeogeographic and temporal uncertainty, dictate the reconstruction of large-scale climate patterns without the option of including additional knowledge (e.g. that latitudinal temperature gradients should be broadly negative).
